# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
setwd("C:/Users/tedco_000/Downloads")
unzip("activity.zip")
d <- read.csv("activity.csv", stringsAsFactors=FALSE, na.strings="NA")
d$date <- as.Date(d$date)
```

## What is mean total number of steps taken per day?

```{r}
library(sqldf)
totstepsday <- sqldf("select date, sum(steps) as sumsteps from d where steps is not null group by 1")
mean(totstepsday$sumsteps)
median(totstepsday$sumsteps)
```

## What is the average daily activity pattern?

```{r}
hist(totstepsday$sumsteps)
avgstepsint <- sqldf("select interval, avg(steps) as avgsteps from d where steps is not null group by 1")
maxavgstepsint <- sqldf("select interval, avgsteps from avgstepsint where avgsteps in (select max(avgsteps) from avgstepsint)")
maxavgstepsint$interval
maxavgstepsint$avgsteps
plot(avgstepsint$interval, avgstepsint$avgsteps, type="l", ylab="Steps", xlab="Interval", main="Average Daily Steps by Interval")
```

## Imputing missing values 
(imputation using mean steps per interval REDUCED mean/median)

```{r}
numNa <- sqldf("select count(*) as numNa from d where steps is null")
numNa
d2 <- sqldf("select coalesce(steps, avgsteps) as steps, date, d.interval from d left join avgstepsint on d.interval=avgstepsint.interval")
totstepsday2 <- sqldf("select date, sum(steps) as sumsteps from d2 where steps is not null group by 1")
mean(totstepsday2$sumsteps)
median(totstepsday2$sumsteps)
hist(totstepsday2$sumsteps)
```

## Are there differences in activity patterns between weekdays and weekends?
(more steps earlier during weekdays - commute to work, perhaps?)
```{r}
d2$daycat <- ifelse((weekdays(d2$date) == "Sunday") | (weekdays(d2$date) == "Saturday"), "Weekend", "Weekday")
avgstepsdaycatintday <- sqldf("select daycat, interval, avg(steps) as avgsteps from d2 where steps is not null group by 1, 2")
library(lattice)
xyplot(avgsteps~interval|daycat, main="Average Steps", data=avgstepsdaycatintday, layout=(c(1,2)), type="l")
```